---
title: "Population Dynamics at HBCUs and Ivy League Institutions"
author: 
  - "Nathan Alexander"
  - "Center for Applied Data Science and Analytics"
  - "Howard University"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true  # Enable floating TOC in the sidebar
    number_sections: true
    theme: cerulean
    always_allow_html: true
    self_contained: true
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remotes::install_github('jbryer/ipeds')
```

```{r ipeds, include=F}
library(ipeds)
library(ggplot2)
library(xtable)
#install.packages("maps")
library(maps)

# list of the currently available functions
ls('package:ipeds')
```

# DATA

Loading data from an [`ipeds` R package](https://github.com/jbryer/ipeds) to analyze population dynamics at HBCUs and Ivy League Institutions.

The Integrated Postsecondary Education Data System, or IPEDS, contains data on higher education institutions in the United States. The data are made publically available for analysis by researchers. 

For this memo, I used the `ipeds` package in R to perform all analysis.


```{r, include=F}
# list of surveys that are available to download
data(surveys)
surveys[,c('SurveyID', 'Title')]

# list of all available CIP codes
data(cipcodes)
names(cipcodes)

# data available
ipeds::available_ipeds() 
```


## Institutions

```{r, eval=F}
ipeds::download_ipeds(
  year = 2018, 
  dir = "/Users/nathanalexander/Dropbox/Work items/Howard/C&I/Memos/data/downloaded",
  force=T) # download data
ipeds2018 <- ipeds::load_ipeds(2018) # load data
names(ipeds2018)
hd2021 <- ipeds::ipeds_survey(table = 'HD', year = 2021)
names(hd2021)
names(hd2021) <- tolower(names(hd2021))
hd2021 <- ipeds::recodeDirectory(hd2021)

HBCUs <- c(
  "220182", # Tennessee State University
  "175772", # Jackson State University
  "234155", # Virginia State University
  "176044", # Mississippi Valley State University
  "198543", # Fayetteville State University
  "198507", # Elizabeth City State University
  "199102", # North Carolina A&T State University
  "131520"  # Howard University
)

hd202.hbcu <- hd2021[which(hd2021$unitid %in%HBCU),]
p <- hd2021.hbcu[, c("instnm", "webaddr", "stabbr", "control")]
names(p) <- c("Institution", "Web Address", "State", "Sector")
p

IvyLeague <- c(
  "186131", # Yale University
  "190150", # Columbia University
  "166027", # Cornell University
  "130794", # Dartmouth College
  "215062", # University of Pennsylvania
  "182670", # Princeton University
  "217156", # Brown University
  "190415"  # Harvard University
)

hd2021.ivy <- hd2021[which(hd2021$unitid %in%IvyLeague),]
q <- hd2021.ivy[, c("instnm", "webaddr", "stabbr", "control")]
names(p) <- c("Institution", "Web Address", "State", "Sector")
q
```
