---
title: "Integrated Postsecondary Education Data System (IPEDS) Data Import"
author:
- Nathan Alexander, PhD
- Amleset Abrhale
- Howard University
output:
  pdf_document:
    toc: true
    toc_depth: '2'
  pdf: null
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: cerulean
    always_allow_html: true
    self_contained: true
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=T, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(stringr)
library(weights)
library(here)
here::i_am("ipeds-import.Rmd")
```

We import and prepare for the analysis of the Integrated Postsecondary Education Data System (IPEDS) surveys. We first review the most recent data sets and dictionaries available (2023) [here](https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx?gotoReportId=7&fromIpeds=true&sid=dd2935f1-8689-48b6-8762-f2258855e466&rtid=1). We also explore some potential options for future analysis.

# DATA

We import a series of raw data files.

- **Institutional Characteristics**: Educational offerings, organization, services and athletic associations

- **Graduation Rates**:	Graduation rate data, 150% of normal time to complete - cohort year 2017 (4-year) and cohort year 2020 (2-year) institutions

- **Student Financial Aid and Net Price**: Student financial aid and net price: 2022-23

```{r}
# institutional characteristics
ic <- read.csv("../data/ipeds-2023-ic.csv")

# institutional characteristics
gradrate <- read.csv("../data/ipeds-2023-grad-rate-4year-cohort17-2year-cohort20.csv")

# institutional characteristics
aid <- read.csv("../data/ipeds-2023-stu-fin-aid-net-price.csv")
```

Then we preview the data.

Given the large size of the data frames, we select the first few columns and limiting to the first two rows.

## Institutional characteristics

```{r, include=T, message=F, warning=F}
ic %>% 
  select(UNITID, PEO1ISTR, PEO2ISTR, PEO3ISTR, PEO4ISTR) %>% 
  head(n=2)
```

## Graduation rate

```{r, include=T, message=F, warning=F}
gradrate %>% 
  select(UNITID, GRTYPE, CHRTSTAT, SECTION, COHORT) %>% 
  head(n=2)
```

## Financial aid

```{r, include=T, message=F, warning=F}
aid %>% 
  select(UNITID, SCUGRAD, SCUGDGSK, SCUGNDGS, SCUGFFN) %>% 
  head(n=2)
```

# VARIABLES

Amleset will work on the variable labels.

# INSTITUTIONS

```{r, eval=F}
hbcu <- c(
  "220182", # Tennessee State University
  "175772", # Jackson State University
  "234155", # Virginia State University
  "176044", # Mississippi Valley State University
  "198543", # Fayetteville State University
  "198507", # Elizabeth City State University
  "199102", # North Carolina A&T State University
  "131520"  # Howard University
)

hd202.hbcu <- hd2021[which(hd2021$unitid %in%hbcu),]
p <- hd2021.hbcu[, c("instnm", "webaddr", "stabbr", "control")]
names(p) <- c("Institution", "Web Address", "State", "Sector")
p

ivy <- c(
  "186131", # Yale University
  "190150", # Columbia University
  "166027", # Cornell University
  "130794", # Dartmouth College
  "215062", # University of Pennsylvania
  "182670", # Princeton University
  "217156", # Brown University
  "190415"  # Harvard University
)

hd2021.ivy <- hd2021[which(hd2021$unitid %in%ivy),]
q <- hd2021.ivy[, c("instnm", "webaddr", "stabbr", "control")]
names(p) <- c("Institution", "Web Address", "State", "Sector")
q
```


